---
title: "Netflix Shows And Movies Analysis"
author: "Caleb Mogyabiyedom"
editor: visual
---


```{r}



suppressPackageStartupMessages(library(tidyverse))
library(ggplot2)
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')



netflix_titles_updated <- netflix_titles |>
  mutate(year_added = str_extract(date_added, "\\d{4}"),
         year_added = str_replace_na(year_added, "Unknown"))|>
  mutate(type = str_to_upper(type))|>
  
  filter(year_added != "Unknown")


#view(netflix_titles_updated)


proportion_data <- netflix_titles_updated |>
  group_by(year_added, type) |>
  summarise(count = n()) |>
  ungroup() |>
  group_by(year_added) |>
  mutate(proportion = count / sum(count))

#proportion_data



proportion_plot_1 <- ggplot(proportion_data, aes(x = year_added, y = proportion, fill = type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Proportion of TV Shows and Movies by Year", 
       x = "Year Added", 
       y = "Proportion", 
       fill = "Type")

proportion_plot_1


```

The first plot, "Proportion of TV Shows and Movies by Year", shows the proportion of TV shows and movies added to Netflix each year. This is a stacked bar chart, where each year is represented by a bar, with segments showing the proportion of TV shows and movies.In the graph, movies tended to dominate the viewing options expecially from 2009 to 2012. Although TV Shows in 2008 had the same as movies. In recent years however,the proportion of TV shows have increased with about 25 percent of viewing options being TV shows.


```{r}
trend_plot <- netflix_titles_updated |>
  count(year_added, type) |>
  ggplot(aes(x = year_added, y = n, color = type, group = type)) +
  geom_line() +
  geom_point() +
  labs(title = "Trend of TV Shows and Movies Added by Year", 
       x = "Year Added", 
       y = "Count", 
       color = "Type")
trend_plot




```
The second plot, "Trend of TV Shows and Movies Added by Year", shows the actual counts of TV shows and movies added per year. This is a line plot with different colors representing movies and TV shows. Between 2015 and 2020, they both experienced steep growths however movies added far outweighed TV Shows added




```{r}
genres <- c("Thriller", "Action", "Dramas", "Romantic", "Horror", "Crime", "Sci-Fi", "Fantasy", "Comedies", "Documentaries", "Kids\' TV")

genre<- netflix_titles_updated|>
  mutate(is_genre = str_detect(listed_in, str_c(genres, collapse = "|")))|>
  mutate(genre_label = str_extract(listed_in, str_c(genres, collapse = "|")))|>
  group_by(year_added, type, genre_label) |>
  summarise(count = n())
#genre

 
ggplot(genre, aes(x = year_added, y = count, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ genre_label) + 
  labs(title = "Variations in Number of Movies and TV Shows by Genre and Year",
       x = "Year Added",
       y = "Count of Titles",
       fill = "Type")+
       theme_minimal()
```
The third plot, "Variations in Number of Movies and TV Shows by Genre and Year", shows how different genres have varied over the years with respect to both TV Shows and Movies. In recent years, there's been growth in all the genres, however Comedies, Dramas, Documentary and Action movies rose more highly. Although Crimes, Kid's TV and Romantic movies have all risen in the TV Shows category, they are still below that of movies.













Reference:

Harmon, J. (n.d.). Tidytuesday/data/2021/2021-04-20/readme.md at main · rfordatascience/tidytuesday. GitHub. https://github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-04-20/readme.md 

The data comes from Kaggle with credit to Shivam Bansal and it was collected from Flixable which is a third-party Netflix search engine.

<https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-20/readme.md>
