---
title: "SQL WAI Analysis"
author: "Caleb Mogyabiyedom"
editor: visual
---



The goal of this analysis is to explore how mean absorbance varies across studies and demographic groups within the Wideband Acoustic Immittance (WAI) database. To begin with, I will I will recreate Figure 1 by analyzing the trends, focusing on mean absorbance values across a specific range of frequencies. 

```{r}
library(RMariaDB)
library(tidyverse)
library(ggplot2)


con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)


figure1_data <- dbGetQuery(con_wai, "
WITH StudyFrequencies AS (
    SELECT 
        m.Identifier,
        p.AuthorsShortList,
        p.Year,
        m.Instrument,
        m.Frequency,
        AVG(m.Absorbance) AS MeanAbsorbance,
        COUNT(DISTINCT m.SubjectNumber, m.Ear) AS UniqueEars
    FROM Measurements m
    JOIN PI_Info p ON m.Identifier = p.Identifier
    WHERE m.Frequency IN (
        210.938, 234.375, 257.812, 281.250, 
        305.688, 330.125, 354.562, 379.000, 
        403.437, 427.875, 452.312, 476.750
    )
    GROUP BY 
        m.Identifier, 
        p.AuthorsShortList,
        p.Year,
        m.Instrument,
        m.Frequency
)

SELECT 
    Identifier,
    AuthorsShortList,
    Year,
    Instrument,
    Frequency,
    MeanAbsorbance,
    UniqueEars
FROM StudyFrequencies
ORDER BY Identifier, Frequency
")

# Create custom legend labels
figure1_data <- figure1_data %>%
  mutate(StudyLabel = paste(
    AuthorsShortList, 
    Year, 
    paste0("N=", UniqueEars), 
    Instrument, 
    sep = ", "
  ))


ggplot(figure1_data, aes(x = Frequency, y = MeanAbsorbance, color = Identifier)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = c(210.938, 281.250, 354.562, 427.875, 476.750),
    labels = c("210", "281", "355", "428", "477")
  ) +
  scale_y_continuous(limits = c(0, 0.5)) +
  labs(
    title = "Mean Absorbances for WAI Studies",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Study"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.key.height = unit(1, "lines")
  ) +
  guides(color = guide_legend(ncol = 1))


dbDisconnect(con_wai)
```
The graph above,illustrates mean absorbances for different WAI studies across a specific frequency range with each line representing a specific study. For most of the studies, there is a slight increase in the mean absorbanses at 281Hz.




For the second aspect of this analysis, I will explore group differences by sex and race, investigating how these demographic variables influence mean absorbance trends. And then, I will calculate mean absorbance for each group across frequencies and visualize these trends with labeled plots.

```{r}
library(DBI)
library(RMariaDB)
library(ggplot2)



con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)


joined_data <- dbGetQuery(con_wai, "
WITH StudyDetails AS (
    SELECT 
        m.Identifier,
        p.AuthorsShortList,
        p.Year,
        s.Sex,
        s.Race,
        m.AgeCategory,
        m.Frequency,
        AVG(m.Absorbance) AS MeanAbsorbance,
        COUNT(DISTINCT m.SubjectNumber) AS NumSubjects,
        COUNT(DISTINCT m.SubjectNumber, m.Ear) AS UniqueEars
    FROM Measurements m
    JOIN Subjects s ON m.SubjectNumber = s.SubjectNumber
    JOIN PI_Info p ON m.Identifier = p.Identifier
    WHERE m.Frequency IN (
        210.938, 234.375, 257.812, 281.250, 
        305.688, 330.125, 354.562, 379.000, 
        403.437, 427.875, 452.312, 476.750
    )
    GROUP BY 
        m.Identifier, 
        p.AuthorsShortList,
        p.Year,
        s.Sex,
        s.Race,
        m.AgeCategory,
        m.Frequency
)

SELECT * FROM StudyDetails
ORDER BY Identifier, Frequency
")


sex_plot <- ggplot(joined_data, aes(x = Frequency, y = MeanAbsorbance, color = Sex)) +
  geom_line(aes(group = interaction(Sex, Identifier))) +
  geom_point() +
  labs(
    title = "Mean Absorbance by Sex Across Frequencies",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Sex"
  ) +
  theme_minimal()


race_plot <- ggplot(joined_data, aes(x = Frequency, y = MeanAbsorbance, color = Race)) +
  geom_line(aes(group = interaction(Race, Identifier))) +
  geom_point() +
  labs(
    title = "Mean Absorbance by Race Across Frequencies",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = "Race"
  ) +
  theme_minimal()


print(sex_plot)
print(race_plot)


dbDisconnect(con_wai)

```

The " Mean Absorbances by Sex Across Frequencies" graph illustrates how mean absorbance varies across frequencies for different sex groups. The variation in absorbance between the sexes is visible across frequencies.

The "Mean Absorbance by Race Across Frequencies" graph also shows the mean absorbance across frequencies for different races.




To recreate Figure 1 from Voss (2020), I queried the WAI database to extract data, filtered for relevant frequencies and combined measurements with authors, year, and instrument. I then calculated the mean absorbance and the number of unique ears  for each study and frequency and then used a line plot to show the relationship. 

To analyze group differences, I examined the impact of sex and race on mean absorbance across frequencies. U I grouped data by study, sex, race, and frequency, calculating mean absorbance for each group. I then generated two plotsto visualize these trends.


DATA SOURCE:https://www.science.smith.edu/wai-database/
